---
title: "Weighting schemes"
output: html_notebook
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(
  message = FALSE,warning = FALSE, fig.width = 6, fig.height = 6)
```

```{r}
#library(e1071)  
library(MCMCpack)
library(tidyverse)
library(ggridges)
library(ggthemes)
library(gridExtra)
set.seed(1234)
```


- The goal is to build a weighted distribution of race by (both) first and last name. 
- If someone has a (first/last) name with a high probability, we want to give it more priority.


# definitions

## Version 1: Standard Deviation or variance 

- For a vector of probabilities with 4 categories, the standard deviation, `sd` moves between `r sd(c(0.25,0.25,0.25,0.25))` for perfect uniformity,  and `r sd(c(0,0,0,1))` for all the probability to one cat.

- the greater the sd, the higher the weight given to the (first/last) name.
- we normalize the values so the sum to 1, with

$$
name_{weight} = \frac{name_{sd}}{name_{sd}+lastname_{sd}}
$$

- We can add a parameter `exp` so we use the sd when `exp=1`, the variance when `exp=2`, and possibly higher orders. The higher the `exp`, the more extreme the weighting scheme 


```{r}
weighting_scheme_sd <- function(name, lastname, exp = 1){
  
  name_sd <- sd(name)^exp
  lastname_sd <- sd(lastname)^exp
  
  total <- name_sd+lastname_sd
  
  if (total==0) {
    name_sd = lastname_sd =1
    total=2
  }
  
  name_weight <- name_sd/total
  lastname_weight <- lastname_sd/total
  
  c('name_weight'=name_weight,'lastname_weight'=lastname_weight)
}
```

## alternative, entropy-based

- A measure that explicitly measure the level of disorder_ in a probability distribution is the concept of entropy.

$$
e= \sum_{i=1}^c -p_i \log_2 p_i
$$

```{r}
entropy <- function(vec){
  vec<-vec[vec>0]
  -sum(vec * log2(vec))}
```

- For a vector of probabilities with 4 categories, the entropy moves between `r entropy(c(0.25,0.25,0.25,0.25))` for perfect uniformity,  and `r entropy(c(0,0,0,1))` for all the probability to one cat.


```{r}
weighting_scheme_entropy <- function(name, lastname, exp=1){
  
  name_e <- entropy(name)^exp
  lastname_e <- entropy(lastname)^exp

  total <- name_e+lastname_e
  
  if (total==0) {
    name_e = lastname_e =1
    total=2
  }
  
  name_weight <- 1- name_e/total
  lastname_weight <- 1- lastname_e/total
  
  c('name_weight'=name_weight,'lastname_weight'=lastname_weight)
}
```



## Simulations generator

for evaluation:

1. generate 150 random distributions for first and last names with `simulate_distributions`
2. calculate weights for each model and exp with `calculate_weights`
3. Build three different plots for analysis with `plot_simulation`  binned by the maximum probability in first/last name (i.e., group by different possible scenarios)

```{r}
simulate_distributions <- function(){
  
  names <- rdirichlet(150,alpha = c(1,1,1,1)) 
  lastnames <- rdirichlet(150,alpha = c(1,1,1,1))
  
  simulations <- expand_grid(name_id=1:150,lastname_id=1:150) %>% 
    mutate(name_dist = map(name_id, ~ names[.x,]),
           lastname_dist = map(lastname_id, ~ lastnames[.x,]),
           name_max =  map(name_dist,max),
           lastname_max = map(lastname_dist,max)
           )
  simulations
}

calculate_weights <- function(simulations,type='sd', exp=1){
  
  
  if (type=='sd') {
    simulations <- simulations %>% 
      mutate(weights = map2(name_dist, lastname_dist,~weighting_scheme_sd(.x,.y,exp)))
  }
  if (type=='entropy') {
    simulations <- simulations %>% 
      mutate(weights = map2(name_dist, lastname_dist,~weighting_scheme_entropy(.x,.y,exp)))
  }
  
  simulations <- simulations %>% 
    unnest(cols = c(name_max,lastname_max,name_max,lastname_max)) %>% 
    arrange(name_max,lastname_max) %>% 
    unnest_wider(weights)
  
  simulations
}

plot_simulation <- function(simulations,title,plot_type='bar'){
  
  
  if (plot_type == 'bar') {
    
    plot <- simulations %>% 
      mutate(name_max = cut(name_max, 10,dig.lab=1),
             lastname_max = cut(lastname_max, 10,dig.lab=1)
      ) %>% 
      group_by(name_max, lastname_max) %>% 
      summarise(name_weight = mean(name_weight),
                lastname_weight = mean(lastname_weight)) %>% 
      pivot_longer(name_weight:lastname_weight) %>% 
      ggplot(aes(x=1,y=value,label=scales::percent(value,2), fill = name, group=name))+
      geom_col(position = 'dodge')+
      geom_text(position = position_dodge(width = 1), size=3,vjust = -0.1)+
      facet_grid(name_max~lastname_max)+
      labs(x='max prop. name',y='max prop. lastname', fill='', title = title)+
      scale_x_continuous(position = "top")+
      scale_y_continuous(position = "right",limits = c(0,1.2))+
      theme(legend.position = 'bottom',
            axis.text = element_blank(),
            axis.ticks = element_blank())
  }
  if (plot_type =='density') {
    
    plot <- simulations %>% 
      mutate(name_max = cut(name_max, 10,dig.lab=1),
             lastname_max = cut(lastname_max, 10,dig.lab=1)
      ) %>% 
      pivot_longer(name_weight:lastname_weight) %>% 
      ggplot(aes(x=value,y=name, fill = name, group=name))+
      geom_density_ridges()+
      facet_grid(name_max~lastname_max)+
      labs(x='max proportion First name',y='max proportion last name', fill='', title = title)+
      #scale_x_continuous(labels = scales::percent)+
      theme(legend.position = 'bottom',
            axis.text.y = element_blank(),
            axis.text.x = element_blank(),
            # axis.text.x = element_text(angle = 90),
            axis.ticks = element_blank(),
            strip.text.x = element_text(angle = 0, size = 6),
            strip.text.y = element_text(angle = 0))+
      scale_fill_gdocs()
  }
  if (plot_type == 'hex'){
    
    plot <- simulations %>% 
      ggplot(aes(x=name_max,y=lastname_max,z=name_weight))+
      stat_summary_hex(fun = function(x) mean(x),bins = 30)+
      labs(x='max prop. name',y='max prop. lastname', fill='Name weight', title = title)+
      scale_fill_continuous(type = 'viridis',labels = scales::percent)+
      theme_minimal()+
      theme(legend.position = 'bottom',
            axis.ticks = element_blank())
    
  }
  plot
}
```


an implementation on how the new weights are calculated
```{r}
weighted_distribution <- function(name, lastname,type='sd', exp=1){
  
  if (type=='sd') {
    weights <- weighting_scheme_sd(name, lastname,exp = exp)
  }
  
  if (type=='entropy') {
    weights <- weighting_scheme_entropy(name, lastname,exp = exp)
  }
  
  
  name*weights['name_weight'] + lastname*weights['lastname_weight']
}
```

# Examples


- extreme case
```{r}
name <- c(1,0,0,0)
lastname <- c(0.25, 0.25, 0.25,0.25)


weighting_scheme_sd(name, lastname) 
weighting_scheme_entropy(name, lastname) 
```

all models give all prob. to name

- even case
```{r}
name <- c(0.25, 0.25, 0.25,0.25)
lastname <- c(0.25, 0.25, 0.25,0.25)

weighting_scheme_sd(name, lastname) 
weighting_scheme_entropy(name, lastname) 
```

all models give equal prob.


- useful name, unuseful lastname

```{r}
name <- c(0.05, 0.05, 0.85,0.05)
lastname <- c(0.3,0.4,.1,.2)
```

```{r}
#with sd
weighting_scheme_sd(name, lastname) 
weighted_distribution(name, lastname,type='sd')
```


```{r}
#with variance
weighting_scheme_sd(name, lastname, exp = 2) 
weighted_distribution(name, lastname,type='sd', exp = 2)
```


```{r}
#with entropy
weighting_scheme_entropy(name, lastname) 
weighted_distribution(name, lastname,type='entropy', exp = 1)

```


```{r}
#with entropy^2
weighting_scheme_entropy(name, lastname, exp = 2) 
weighted_distribution(name, lastname,type='entropy', exp = 2)
```


# Simulations

```{r}

simulated_names <- simulate_distributions()

weights_sd <- calculate_weights(simulated_names, type = 'sd',exp = 1)
weights_var <- calculate_weights(simulated_names, type = 'sd',exp = 2)

weights_entropy <- calculate_weights(simulated_names, type = 'entropy',exp = 1)
weights_entropy2 <- calculate_weights(simulated_names, type = 'entropy',exp = 2)
```

```{r}
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}
```


## Bar plots

```{r}
p1 <- plot_simulation(weights_sd,plot_type='bar',title = 'Standard Deviation')
p2 <- plot_simulation(weights_var,plot_type='bar',title = 'Variance')
p3 <- plot_simulation(weights_entropy,plot_type='bar',title = 'Entropy')
p4 <- plot_simulation(weights_entropy2,plot_type='bar',title = 'Entropy^2')

grid.arrange(p1,p2,p3,p4, nrow=2)
```

## Density plots

```{r message=FALSE, warning=FALSE}
p1 <- plot_simulation(weights_sd,plot_type='density',title = 'Standard Deviation') + 
  theme(legend.position="bottom") +
  labs(x='')
p2 <- plot_simulation(weights_var,plot_type='density',title = 'Variance') + 
  theme(legend.position="none") +
  labs(x='', y ='')
p3 <- plot_simulation(weights_entropy,plot_type='density',title = 'Entropy')  + 
  theme(legend.position="none")
p4 <- plot_simulation(weights_entropy2,plot_type='density',title = 'Entropy^2')  + 
  theme(legend.position="none")+
  labs(y='')


mylegend<-g_legend(p1)

density_plot <- grid.arrange(arrangeGrob(p1 + theme(legend.position="none"),
                         p2, p3,p4, nrow=2),
             mylegend, nrow=2,heights=c(10, 1))

ggsave('../results/plots/density_plot_simulation.png', plot = density_plot, width = 12, height = 6, dpi = 300)

```


## Hex plots


```{r}

weights_sd %>% 
  mutate(type = 'Standard Deviation') %>% 
  bind_rows(weights_var %>% 
              mutate(type = 'Variance')) %>% 
  bind_rows(weights_entropy %>% 
              mutate(type = 'Entropy')) %>%
  bind_rows(weights_entropy2 %>% 
              mutate(type = 'Entropy^2')) %>% 
  ggplot(aes(x=name_max,y=lastname_max,z=name_weight))+
  stat_summary_hex(fun = function(x) mean(x),bins = 30)+
  labs(x='max proportion first name',y='max proportion last name', fill='First Name weight')+
  facet_wrap(.~ type)+
  scale_fill_continuous(type = 'viridis',labels = scales::percent)+
  theme_minimal()+
  scale_x_continuous(labels = scales::percent)+
  scale_y_continuous(labels = scales::percent)+
  theme(axis.ticks = element_blank(),
        text = element_text(size = 18))
ggsave('../results/plots/hex_plot_simulation.png', width = 12, height = 6, dpi = 300)
```


